Four reasons your build is slow

### 2019-01-17 paper.tex
edits

In this section we discuss four problems with project builds that use make.  In section \ref{sec:expensive},
we discuss
builds whose spans are dominated by a single expensive command that prevents the build from
achieving greater parallelism, look at a project that exhibits this behavior, and explain our
our tool can diagnose this problem and identify areas for improvement.  In section \ref{sec:sequential}
, we discuss what
it means for a build to be unnecessarily sequential, and look at an example of a build that
exhibits this behavior, and how our tool can diagnose this problem and identify areas for improvement.
In section \ref{sec:scalability}, we discuss when a build should have ample available parallelism
but fails to achieve that
parallelism in practice, look at a project that exhibits this behavior, and discuss potential causes.
In section \ref{sec:compiler} we discuss another instance of an expensive command dominating a
build's span; in this case a
slow compiler, we look at a project that exhibits this behavior and how our tool can diagnose it
and identify areas for improvement.

@@

To diagnose the problem of a span dominated by a single long running task, our tool can be used
to learn the span path through the build graph and learn which target and shell commands
are the most time intensive.  To see what sort of an improvement would occur in the span if the
running time of the most expensive target in the span was to decrease, the tool will recalculate
the span of the graph excluding its most expensive member any number of times.  This is useful
in the case where decreasing the running time of a single task gives the build graph a new span
path with a similarly long running time, which is dominated by a different long running command.

Facebook's zstd project \cite{zstd1.3.7} has a build that exhibits this property.  When we used our tool to analyze
the span of zstd's build we saw that it is dominated by a single call to cc; see fig. \ref{code:cc1}.

\begin{figure}[H]
\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small]
  cc \$(FLAGS) common/debug.c \textrm{\emph{26 more .c files}}
    zstdcli.o fileio.o bench.o datagen.o dibio.o
    -o zstd
\end{Verbatim}
\caption{A long running cc command which was responsible for about 97\% of zstd's span.  It compiled and linked 27 .c files.}
\label{code:cc1}
\end{figure}

If this call is no longer part of the span, the new span is similarly dominated by another single call to cc;
see fig. \ref{code:cc2}.
\begin{figure}[H]
\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small]
  cc \$(FLAGS) common/debug.c \textrm{\emph{29 more .c files}}
    -o libzstd.so.1.3.7
\end{Verbatim}
\caption{A long running cc command which was responsible for approximately 100\% of zstd's span when the command in fig. \ref{code:cc1} is removed.
This command compiled and linked 30 files.}
\label{code:cc2}
\end{figure}
If neither of these calls are members of the build's span, then our tool predicts the span would be
approximately 7.2 seconds rather than 39 seconds, as with the original span, or 37 seconds, as with the second span.
This is about an 80\% reduction in the build's span, and the parallelism improves from 2.98 to 16.

To test this we manually replaced the cc command in fig. \ref{code:cc1} with the code in fig.
\ref{code:cc3}.  The original command is replace by a new target for each object file and target
that can build each object file in parallel and link them together afterwards.

@@

Figure \ref{fig:zstd} shows that when the modified build was run with 1 to 32 jobs,
the predicted speedup was observed.  An additional performance improvement that was not predicted
is that the work of the new build decreased from approximately 120 seconds to approximately 80 seconds.
This is because each of the object files is declared as its own target, so make can consider whether
or not the file already exists and if it is necessary to rebuild it.  In this case, it turns out that
the object files being built as part of the cc command in fig. \ref{code:cc1} had already been built
by make; making it unnecessary to build them again.  This cuts the effect time of the code in fig.
\ref{code:cc1} from about 40 seconds to about 2 seconds; just the time to link the object files.

\subsection{Unnecessarily sequential}
\label{sec:sequential}

>>>>

In this section we discuss four problems with project builds that use
make.  In section \ref{sec:expensive}, we discuss builds whose spans
are dominated by a single expensive command that prevents the build
from achieving greater parallelism, look at a project that exhibits
this behavior, and explain our our tool can diagnose this problem and
identify areas for improvement.  In section \ref{sec:sequential} , we
discuss what it means for a build to be unnecessarily sequential, and
look at an example of a build that exhibits this behavior, and how our
tool can diagnose this problem and identify areas for improvement.  In
section \ref{sec:scalability}, we discuss when a build should have
ample available parallelism but fails to achieve that parallelism in
practice, look at a project that exhibits this behavior, and discuss
potential causes.  In section \ref{sec:compiler} we discuss another
instance of an expensive command dominating a build's span; in this
case a slow compiler, we look at a project that exhibits this behavior
and how our tool can automatically diagnose it and identify areas for improvement.

@@

To diagnose the problem of a span dominated by a single long running
task, our tool can be used to learn the span path through the build
graph and learn which target and shell commands are the most time
intensive.  To see what sort of an improvement would occur in the span
if the running time of the most expensive target in the span was to
decrease, the tool will recalculate the span of the graph excluding
its most expensive member any number of times.  This is useful in the
case where decreasing the running time of a single task gives the
build graph a new span path with a similarly long running time, but
 dominated by a \emph{different} long running command.

Facebook's zstd project~\cite{zstd1.3.7} has a build that exhibits
this property.  When we used our tool to analyze the span of zstd's
build we saw that it is dominated by a single call to \verb|cc|; see
fig.~\ref{code:cc1}.

\begin{figure}[H]
\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small]
  cc \$(FLAGS) common/debug.c \textrm{\emph{26 more .c files}}
    zstdcli.o fileio.o bench.o datagen.o dibio.o
    -o zstd
\end{Verbatim}
\caption{This is responsible for about 97\% of zstd's span.  It compiles and links 27 .c files.}
\label{code:cc1}
\end{figure}

If this command is removed from the build, the new span is similarly dominated by another single call to \verb|cc|; see fig. \ref{code:cc2}.
\begin{figure}[H]
\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small]
  cc \$(FLAGS) common/debug.c \textrm{\emph{29 more .c files}}
    -o libzstd.so.1.3.7
\end{Verbatim}
\caption{A long running cc command responsible for approximately 100\%
  of zstd's span when the command in fig. \ref{code:cc1} is removed.
  This command compiles and links 30 files.}
\label{code:cc2}
\end{figure}
If neither of these calls are members of the build's span, then our
tool predicts the span would be approximately 7.2 seconds rather than
39 seconds, as with the original span, or 37 seconds, as with the
second span.  This is an 80\% reduction in the build's span, and
the parallelism improves from 2.98 to 16.

To test this we manually replaced the cc command in
fig. \ref{code:cc1} with the code in fig.  \ref{code:cc3}.  The
original command is replaced by a new target for each object file and
target that can build each object file in parallel and link them
together afterwards.

@@

Figure \ref{fig:zstd} shows that when the modified build was run with
1 to 32 jobs, the predicted speedup was observed.  An additional
performance improvement that was not predicted is that the work of the
new build decreased from approximately 120 seconds to approximately 80
seconds.  This is because each of the object files is declared as its
own target, so \verb|make| can consider whether or not the file
already exists and if it is necessary to rebuild it.  In this case, it
turns out that the object files being built as part of the cc command
in fig. \ref{code:cc1} had already been built; making it unnecessary
to build them again.  This cuts the effective time of the code in fig.
\ref{code:cc1} from 40 seconds to 2 seconds---just the time to link
the object files.

\subsection{Unnecessarily sequential rules}
\label{sec:sequential}
